---
title: "tidypredict"
output: github_document
---

```{r setup, include=FALSE}
library(dplyr)
library(tidypredict)
library(randomForest)
```

Run predictions inside the database. It works with several databases back-ends because it uses `dbplyr` for the final SQL translation of the algorithm.

## Installation

Install `tidypredict` using `devtools` as follows:
```{r, eval = FALSE}
devtools::install_github("edgararuiz/tidypredict")
```


## Intro

`tidypredict` is able to parse an R model object, such as:
```{r}
model <- lm(mpg ~ wt + cyl, data = mtcars)
```

And then creates the SQL statement needed to calculate the fitted prediction:

```{r}
tidypredict_sql(model, dbplyr::simulate_mssql())
```

## Supported models

The following R models are currently supported. For more info please review the corresponding vignette:

- [Linear Regression](http://tidypredict.netlify.com/vignettes/lm/) - `lm()`
- [Generalized Linear model](http://tidypredict.netlify.com/vignettes/glm/) - `glm()`
- [Random Forest](http://tidypredict.netlify.com/vignettes/randomforest/) - `randomForest()`


