---
title: "Oblique Random Forest, using aorsf"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Oblique Random Forest, using aorsf}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

library(dplyr)
library(tidypredict)
library(parsnip)
library(aorsf)
set.seed(100)
```

| Function                                                      |Works|
|---------------------------------------------------------------|-----|
|`tidypredict_fit()`, `tidypredict_sql()`, `parse_model()`      |  ✔  |
|`tidypredict_to_column()`                                      |  ✗  |
|`tidypredict_test()`                                           |  ✔  |
|`tidypredict_interval()`, `tidypredict_sql_interval()`         |  ✗  |
|`parsnip`                                                      |  ✔  |

Only regression models with numeric predictors are supported. Classification requires a voting mechanism that cannot be expressed as a single formula, and oblique splits on categorical predictors operate on an internal encoding that is not reproduced here.

## How it works

Here is a simple `orsf()` model using the `mtcars` dataset:
```{r}
library(dplyr)
library(tidypredict)
library(aorsf)

model <- orsf(mtcars, mpg ~ ., n_tree = 5)
```

## Under the hood

Unlike axis-aligned forests, each split in an oblique random forest is a linear combination of (standardized) predictors compared against a cutpoint. `tidypredict` reads the trees stored in the fitted forest, folds the forest's centering and scaling into the split coefficients, and turns each tree into a nested `dplyr::case_when()` statement. The trees are then averaged.

```{r}
tidypredict_fit(model)
```

From there, the Tidy Eval formula can be used anywhere where it can be operated. `tidypredict` provides three paths:

  - Use directly inside `dplyr`,  `mutate(mtcars, !! tidypredict_fit(model))`
  - Use `tidypredict_to_column(model)` to a piped command set
  - Use `tidypredict_sql(model)` to retrieve the SQL statement

## A note on split boundaries

`aorsf` uses observed linear-combination values from the training data as split cutpoints. A training row can therefore land exactly on a split boundary, where floating-point differences between `aorsf`'s internal traversal and the generated formula may send it down a different branch. This affects only rows that coincide with a training cutpoint; on new data the formula reproduces `predict()` exactly.

## parsnip

`tidypredict` also supports `aorsf` model objects fitted via the `parsnip` package (using the `bonsai` extension).

```{r}
library(parsnip)
library(bonsai)

parsnip_model <- rand_forest(mode = "regression", trees = 5) %>%
  set_engine("aorsf") %>%
  fit(mpg ~ ., data = mtcars)

tidypredict_fit(parsnip_model)
```
