---
title: "bart models"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{bart models}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
if (requireNamespace("dbarts", quietly = TRUE)) {
  library(tidypredict)
  library(dplyr)
  eval_code <- TRUE
} else {
  eval_code <- FALSE
}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = eval_code
)
```

| Function                                                      |Works|
|---------------------------------------------------------------|-----|
|`tidypredict_fit()`, `tidypredict_sql()`, `parse_model()`      |  ✔  |
|`tidypredict_to_column()`                                      |  ✔  |
|`tidypredict_test()`                                           |  ✔  |
|`tidypredict_interval()`, `tidypredict_sql_interval()`         |  ✗  |
|`parsnip`                                                      |  ✔  |

`dbarts::bart()` fits a Bayesian additive regression trees (BART) model. Each
draw of the posterior sampler holds a full ensemble of trees, and the fitted
value is the average, over every draw, of the summed leaf values of that draw's
trees. `tidypredict_fit()` returns one nested `case_when()` per tree of per
draw, so the size of the returned expression grows with both the number of trees
and the number of posterior samples that were kept. Small values of `ntree` and
`ndpost` keep the formula manageable.

Two things are needed for the model to be parsed:

- The model has to be fit with `keeptrees = TRUE`, otherwise the trees are
  discarded and cannot be recovered.
- The outcome has to be continuous. Binary outcomes are fit with a probit link,
  which cannot be translated to SQL.

## `tidypredict_` functions

```{r}
set.seed(100)
model <- dbarts::bart(
  mtcars[c("wt", "cyl", "disp")],
  mtcars$mpg,
  ntree = 5,
  ndpost = 5,
  keeptrees = TRUE,
  verbose = FALSE
)
```

- Create the R formula
    ```{r}
tidypredict_fit(model)
    ```

- Add the predictions to the original table
    ```{r}
mtcars %>%
  tidypredict_to_column(model) %>%
  glimpse()
    ```

- Confirm that the results match the model's `predict()` results
    ```{r}
tidypredict_test(model, mtcars)
    ```

- Get the SQL translation
    ```{r}
tidypredict_sql(model, dbplyr::simulate_mssql())
    ```

## parsnip

Models fit with `parsnip::bart()` and the `"dbarts"` engine are supported as
well. Note that `predict()` on a parsnip `bart()` model draws from the posterior
*predictive* distribution, which adds residual noise, so its results are only
equal to the ones of `tidypredict_fit()` in expectation.

```{r}
library(parsnip)

set.seed(100)
model <- bart(mode = "regression", trees = 5) %>%
  set_engine("dbarts", ndpost = 5, verbose = FALSE) %>%
  fit(mpg ~ wt + cyl + disp, data = mtcars)

tidypredict_fit(model)
```

## Factor predictors

Factor and character predictors are expanded into indicator columns before the
model is fit. `tidypredict_fit()` maps those columns back onto the original
columns, so the returned formula can be used with the same data that the model
was fit on.

```{r}
set.seed(100)
model <- dbarts::bart(
  data.frame(wt = mtcars$wt, cyl = factor(mtcars$cyl)),
  mtcars$mpg,
  ntree = 2,
  ndpost = 2,
  keeptrees = TRUE,
  verbose = FALSE
)

tidypredict_fit(model)
```
