---
title: "Gradient boosting, using H2O"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Gradient boosting, using H2O}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
# H2O predictions require a running H2O cluster, so the chunks below are only
# evaluated when h2o is installed and a cluster can be started.
h2o_available <- FALSE
if (requireNamespace("h2o", quietly = TRUE)) {
  h2o_available <- tryCatch(
    {
      suppressMessages(h2o::h2o.init())
      h2o::h2o.no_progress()
      TRUE
    },
    error = function(e) FALSE
  )
}

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = h2o_available
)

library(dplyr)
library(tidypredict)
set.seed(100)
```

| Function                                                      |Works|
|---------------------------------------------------------------|-----|
|`tidypredict_fit()`, `tidypredict_sql()`                       |
|`tidypredict_test()`                                           |
|`parsnip`                                                      |

## How it works

`tidypredict` supports H2O gradient boosting models (GBM) fitted through the
`parsnip` package with the `"h2o_gbm"` engine of `boost_tree()`, which is
provided by the `agua` package. Both regression and classification are
supported.

```{r}
library(parsnip)
library(agua)

model <- boost_tree(mode = "regression", trees = 10) |>
  set_engine("h2o_gbm") |>
  fit(mpg ~ wt + cyl + hp, data = mtcars)
```

## Under the hood

An H2O model is a handle into a running H2O cluster. `tidypredict` pulls each
boosted tree from the cluster with `h2o::h2o.getModelTree()` and turns it into a
nested `dplyr::case_when()` statement. The individual tree contributions are
added together (along with the model's initial value) to form the Tidy Eval
formula. Because the trees live in the cluster, an active `h2o.init()`
connection is required.

```{r}
tidypredict_fit(model)
```

From there, the Tidy Eval formula can be used anywhere it can be operated.
`tidypredict` provides these paths:

- Use directly inside `dplyr`, `mutate(mtcars, !! tidypredict_fit(model))`
- Use `tidypredict_sql(model)` to retrieve the SQL statement

```{r}
tidypredict_sql(model, dbplyr::simulate_dbi())
```

## Categorical predictors

H2O handles categorical predictors natively. The generated formula uses `%in%`
for categorical splits.

```{r}
mtcars2 <- mtcars
mtcars2$cyl <- factor(mtcars2$cyl)
mtcars2$gear <- factor(mtcars2$gear)

model_cat <- boost_tree(mode = "regression", trees = 10) |>
  set_engine("h2o_gbm") |>
  fit(mpg ~ cyl + gear + wt, data = mtcars2)

tidypredict_fit(model_cat)
```

## Classification

For binary classification, `tidypredict_fit()` returns the probability of the
second (positive) outcome level via the logistic link.

```{r}
mtcars3 <- mtcars
mtcars3$vs <- factor(mtcars3$vs)

model_bin <- boost_tree(mode = "classification", trees = 10) |>
  set_engine("h2o_gbm") |>
  fit(vs ~ wt + cyl + hp, data = mtcars3)

tidypredict_fit(model_bin)
```

For multiclass classification, `tidypredict_fit()` returns a named list with one
softmax probability formula per class.

```{r}
model_multi <- boost_tree(mode = "classification", trees = 10) |>
  set_engine("h2o_gbm") |>
  fit(Species ~ ., data = iris)

names(tidypredict_fit(model_multi))
```

## Limitations

Only H2O's GBM models are supported, not H2O's XGBoost. The gaussian,
bernoulli, and multinomial distributions are supported. Because predictions
require a live H2O cluster, the parsed formula cannot be reused after the
cluster is shut down.
