---
title: "mixOmics PLS models"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{mixOmics PLS models}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
if (requireNamespace("mixOmics", quietly = TRUE)) {
  library(tidypredict)
  library(dplyr)
  eval_code <- TRUE
} else {
  eval_code <- FALSE
}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = eval_code
)
```

| Function                                                      |Works|
|---------------------------------------------------------------|-----|
|`tidypredict_fit()`, `tidypredict_sql()`, `parse_model()`      |  ✔  |
|`tidypredict_to_column()`                                      |  ✔* |
|`tidypredict_test()`                                           |  ✔* |
|`tidypredict_interval()`, `tidypredict_sql_interval()`         |  ✗  |
|`parsnip`                                                      |  ✔  |

\* Only for regression models with a single outcome.

`mixOmics` fits partial least squares models with `pls()` and `spls()`, and the
discriminant analysis variants with `plsda()` and `splsda()`. All four collapse
down to one set of regression coefficients per outcome, so the generated
expression is a plain linear combination of the predictors.

For a single-outcome regression model, `tidypredict_fit()` returns one
expression. For a model with multiple outcome columns it returns a *named list*
of expressions, one per outcome. For the discriminant variants it returns a
named list of class-probability expressions (the softmax of the predicted dummy
outcomes, matching what `plsmod` predicts). Since those two cases return a list,
`tidypredict_to_column()` and `tidypredict_test()` only work for single-outcome
regression models.

The number of components (`ncomp`) is baked into the coefficients, and the
sparse variants simply give the predictors that were not selected a coefficient
of zero.

## `tidypredict_` functions

```{r}
x <- as.matrix(mtcars[c("cyl", "disp", "hp", "drat")])
model <- mixOmics::pls(x, mtcars$mpg, ncomp = 2)
```

- Create the R formula
    ```{r}
tidypredict_fit(model)
    ```

- Add the prediction to the original table
    ```{r}
library(dplyr)

mtcars %>%
  tidypredict_to_column(model) %>%
  glimpse()
    ```

- Confirm that the results match the model's `predict()` results
    ```{r}
tidypredict_test(model, mtcars)
    ```

## Discriminant analysis

`plsda()` and `splsda()` return one probability expression per class:

```{r}
da_model <- mixOmics::splsda(as.matrix(iris[1:4]), iris$Species, ncomp = 2)

fit <- tidypredict_fit(da_model)
names(fit)
fit[["setosa"]]
```

## parsnip

`parsnip` fitted models are also supported by `tidypredict`. The `mixOmics`
engine of `pls()` works for both regression and classification:

```{r}
library(parsnip)
library(plsmod)

p_model <- pls(num_comp = 2) %>%
  set_engine("mixOmics") %>%
  set_mode("regression") %>%
  fit(mpg ~ disp + hp + drat, data = mtcars)

tidypredict_fit(p_model)
```

`mixOmics` models are fit from a numeric matrix, so a model fit directly can
only refer to the matrix columns it was given. Categorical predictors therefore
work through the `parsnip` interface, which keeps the formula around and lets the
dummy columns be written in terms of the original factors:

```{r}
cars <- transform(mtcars, gear = factor(gear))

c_model <- pls(num_comp = 2) %>%
  set_engine("mixOmics") %>%
  set_mode("regression") %>%
  fit(mpg ~ disp + hp + gear, data = cars)

tidypredict_fit(c_model)
```

## Parse model spec

Here is an example of the model spec:
```{r}
pm <- parse_model(model)
str(pm, 2)
```
