---
title: "Non-R Models"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{non-r}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

A model that is trained in any language are able to integrate with `tidypredict`, and thus with `broom`.  The requirement is that the model in that language is exported using the parse model spec.  The easiest file format would be YAML. 

## python example

A model that was fitted using `sklearn`'s `linear_model`. The model is based on diabetes data. Ten baseline variables, age, sex, body mass index, average blood pressure, and six blood serum measurements were obtained for each of n = 442 diabetes patients, as well as the response of interest, a quantitative measure of disease progression one year after baseline. The model's results were converted to YAML by the same python script, I copied and pasted the top part here:

```
general:
  is_glm: 0
  model: lm
  residual: 0
  sigma2: 0
  type: regression
  version: 2.0
terms:
- coef: 152.76430691633442
  fields:
  - col: (Intercept)
    type: ordinary
  is_intercept: 1
  label: (Intercept)
```

## Read in R

The YAML data can be read in R by using the `yaml` package.  In this example, we have copy-pasted most of the models inside a variable called `sklearn_model`. Because `yaml` requires local YAML variables to be split by line, we use `strsplit()`.

```{r}
library(yaml)

sklearn_model <- strsplit("general:
  is_glm: 0
  model: lm
  residual: 0
  sigma2: 0
  type: regression
  version: 2.0
terms:
- coef: 152.76430691633442
  fields:
  - col: (Intercept)
    type: ordinary
  is_intercept: 1
  label: (Intercept)
- coef: 0.3034995490660432
  fields:
  - col: age
    type: ordinary
  is_intercept: 0
  label: age
- coef: -237.63931533353403
  fields:
  - col: sex
    type: ordinary
  is_intercept: 0
  label: sex
- coef: 510.5306054362253
  fields:
  - col: bmi
    type: ordinary
  is_intercept: 0
  label: bmi
- coef: 327.7369804093466
  fields:
  - col: bp
    type: ordinary
  is_intercept: 0
  label: bp
- coef: -814.1317093725387
  fields:
  - col: s1
    type: ordinary
  is_intercept: 0
  label: s1
", split = "\n")[[1]]
```

Now the model is converted to an R `list` using `yaml.load`.
```{r}
sklearn_model <- yaml.load(sklearn_model)

str(sklearn_model, 2)
```

## `tidypredict`

The `list` object needs to be recognized as a `tidypredict` parsed model.  To do that, we use `as_parsed_model()`

```{r}
library(tidypredict)

spm <- as_parsed_model(sklearn_model)

class(spm)
```

The `spm` variable now works just as any parsed model inside R.  Use `tidypredict_fit()` to view the resulting formula.

```{r}
tidypredict_fit(spm)
```

Now, the model can run **inside a database**

```{r}
tidypredict_sql(spm, dbplyr::simulate_mssql())
```

## `broom`

Now that we have a `parsed_model` object, it is possible to use `broom`'s `tidy()` function.  This means that we are able to integrate a totally external model, with `broom`.

```{r}
tidy(spm)
```

