---
title: "Random Forest, using Ranger"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ranger}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

library(dplyr)
library(tidypredict)
library(tidymodels)
library(ranger)
set.seed(100)
```

## Highlights & Limitations

- Uses the `ranger::getTree()` to parse each decision path.
- In-line functions in the formulas are **not supported**:  
     - OK - `wt ~ mpg + am` 
     - OK - `mutate(mtcars, newam = paste0(am))` and then `wt ~ mpg + newam`
     - Not OK - `wt ~ mpg + as.factor(am)`
     - Not OK - `wt ~ mpg + as.character(am)`
- Interval functions are not supported: `tidypredict_interval()` & `tidypredict_sql_interval()`
   

## How it works

Here is a simple `ranger()` model using the `iris` dataset:
```{r}
library(dplyr)
library(tidypredict)
library(ranger)

model <- ranger(Species ~ .,data = iris, num.trees =  100)
```

## Under the hood

The parser is based on the output from the `ranger::treeInfo()` function.  It will return as many decision paths as there are non-NA rows in the `prediction` field.
```{r}
treeInfo(model) %>%
  head()
```


The output from `parse_model()` is transformed into a `dplyr`, a.k.a Tidy Eval, formula.  The entire decision tree becomes one `dplyr::case_when()` statement
```{r}
tidypredict_fit(model)[1]
```

From there, the Tidy Eval formula can be used anywhere where it can be operated. `tidypredict` provides three paths:

  - Use directly inside `dplyr`,  `mutate(iris, !! tidypredict_fit(model))`
  - Use `tidypredict_to_column(model)` to a piped command set
  - Use `tidypredict_to_sql(model)` to retrieve the SQL statement


## parsnip

`tidypredict` also supports `ranger` model objects fitted via the `parsnip` package.

```{r}
library(tidymodels)

parsnip_model <- rand_forest() %>%
  set_engine("ranger") %>%
  fit(Species ~ ., data = iris)

tidypredict_fit(parsnip_model)[[1]]
```