---
title: "Random Forest"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{rf}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

library(dplyr)
library(tidypredict)
library(randomForest)
set.seed(100)
```

## Highlights & Limitations

- Uses the `randomForest::getTree()` to parse each decision path.
- In-line functions in the formulas are **not supported**:  
     - OK - `wt ~ mpg + am` 
     - OK - `mutate(mtcars, newam = paste0(am))` and then `wt ~ mpg + newam`
     - Not OK - `wt ~ mpg + as.factor(am)`
     - Not OK - `wt ~ mpg + as.character(am)`
- Interval functions are not supported: `tidypredict_interval()` & `tidypredict_sql_interval()`
   

## How it works

Here is a simple `randomForest()` model using the `iris` dataset:
```{r}
library(dplyr)
library(tidypredict)
library(randomForest)

model <- randomForest(Species ~ .,data = iris ,ntree = 100, proximity = TRUE)
```

## Under the hood

The parser is based on the output from the `randomForest::getTree()` function.  It will return as many decision paths as there are non-NA rows in the `prediction` field.
```{r}
getTree(model, labelVar = TRUE) %>%
  head()
```


The output from `parse_model()` is transformed into a `dplyr`, a.k.a Tidy Eval, formula.  The entire decision tree becomes one `dplyr::case_when()` statement
```{r}
tidypredict_fit(model)[1]
```

From there, the Tidy Eval formula can be used anywhere where it can be operated. `tidypredict` provides three paths:

  - Use directly inside `dplyr`,  `mutate(iris, !! tidypredict_fit(model))`
  - Use `tidypredict_to_column(model)` to a piped command set
  - Use `tidypredict_to_sql(model)` to retrieve the SQL statement
